---
title: "Sourcing Data on Internet Access in USA"
author: "Alissa Stover"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose 

This program loads data from the Federal Communications Commission on internet access in the USA. The data was manually downloaded from this URL: https://www.fcc.gov/general/form-477-county-data-internet-access-services 

The data "Includes data on residential, non-residential, and total fixed Internet access connections over 200 kbps in at least one direction by county, and county households." 

These data are exported to a folder containing an R Shiny app so that they can be used directly for the app. 

## Prepare environment

### Load packages

```{r}
# set directory & file name 
dir <- "~/Documents/data_science/rshiny_tech-access/source_data"
# run script to load saved file paths
source(file.path(dir, "configs.R"))
```

```{r, message=FALSE}
# run script to load packages 
source(file.path(dir, "packages.R"))
```

### Load data

```{r}
# load data dictionary - on sheet 1
data_dict <- read_excel(file.path(dir, raw_data), sheet = 1)
```

```{r}
# show data dictionary 
data_dict %>%
  knitr::kable()
```



```{r}
# load data - on sheet 2
# "-9999" are coded as NAs 
# from the data dictionary, we know that -9999 means data withheld to maintain firm confidentiality
data <- readxl::read_excel(file.path(dir, raw_data), sheet = 2, na = "-9999")
```

## Inspect data 

### Visual check of first 5 rows 

```{r}
# print first 5 rows 
head(data) %>%
  knitr::kable()
```

### Check datatypes

We can see that we have 3,234 rows and 8 columns, which are as expected according to the data dictionary. 

```{r}
# print data type and first few values for each column 
dplyr::glimpse(data) 
```

#### All variables appear to be the appropriate data type 

The 8 variables along with their meaning and datatype are as follows: 

* `countycode`: County FIPS code; string
* `statename`: State name (complete); string
* `countyname`: County name (complete); string
* `consumer`: 	Residential Fixed High-Speed Connections over 200 kbps in at least one direction, in thousands; numeric (double)
* `non_consumer`: Non-residential Fixed High-Speed Connections over 200 kbps in at least one direction, in thousands; numeric (double)
* `all`: Total Fixed High-Speed Connections over 200 kbps in at least one direction, in thousands; numeric (double)
* `hhs`: Households, in thousands (Census: 2013-2017 5-year ACS); numeric (double)
* `ratio`: 	residential / hhs; numeric (double)

## Merge to other datasets 

One useful thing to with these data is plot them on a map. The `urbnmapr`package has useful helper functions for this, which require you to merge your data onto their dataset with geometric information. 

```{r}
# county-level data
urbn_map_county <- urbnmapr::get_urbn_map(map = "counties", sf = TRUE) %>%
  left_join(data, by = c("county_name" = "countyname"))
# state-level data
urbn_map_state <- urbnmapr::get_urbn_map(map = "states", sf = TRUE) %>%
  left_join(data, by = c("state_name" = "statename"))

```
## Save data

```{r}
# save as RData objects
save(data, file = file.path(dir, rda_data))
save(urbn_map_county, file = file.path(dir, "../tech_access_app/urbn_map_county.rda"))
save(urbn_map_state, file = file.path(dir, "../tech_access_app/urbn_map_state.rda"))
```
